---
title: "Mote BCRS Data"
author: "Lance Price"
date: "June 22, 2017"
output: pdf_document
---

# Import data

```{r}
library(openxlsx)

# Directory Path
mote_dir <- "C:/Users/lance/Desktop/GIS/Mote_Marine_Lab/"

# File Path
bcrs_file <- paste(mote_dir, "/Data/BCRS_2012to2017.xlsx", sep = "")

# Import BCRS data
bcrs_2012_2015 <- read.xlsx(bcrs_file, sheet = 1, rows = seq(1,33596))
bcrs_2015_2017 <- read.xlsx(bcrs_file, sheet = 2, na.strings = "NULL", cols = c(seq(1,20),seq(22,31)))

# Date times were changed to numeric values on import
# Convert them back to date times
bcrs_2012_2015$date <- convertToDateTime(bcrs_2012_2015$date, origin = '1900-01-01')
bcrs_2015_2017$created_at <- convertToDateTime(bcrs_2015_2017$created_at, origin = '1900-01-01')
bcrs_2015_2017$updated_at <- convertToDateTime(bcrs_2015_2017$updated_at, origin = '1900-01-01')
```


# Clean data

```{r}
library(dplyr)

# In 2012-2015 data, flags that had no value were incorrectly labeled with the county name
# I correct this
mislabeled <- which(is.na(bcrs_2012_2015$county))
mislabeled_flags <- bcrs_2012_2015$flag[mislabeled]
bcrs_2012_2015$flag[mislabeled] <- NA
bcrs_2012_2015$county[mislabeled] <- mislabeled_flags

# In 2015-2017 data, surface and air temperatures have 0 values for missing data.
# Change them to NA
bcrs_2015_2017$water_surface_tempature[which(bcrs_2015_2017$water_surface_tempature == 0)] <- NA
bcrs_2015_2017$air_tempature[which(bcrs_2015_2017$air_tempature == 0)] <- NA

# make data frame of beach info
beaches <- bcrs_2015_2017[1:33595,c(4,7,8)]
beaches <- unique(beaches)

bcrs_2015_2017 <- bcrs_2015_2017[, -c(7,8)]

bcrs_2015_2017 <- inner_join(bcrs_2015_2017,beaches, by="beach_id")

write.csv(bcrs_2015_2017, file = paste(mote_dir,"/Data/BCRS_2015to2017LatLong.csv", sep = ""))
```

# Put 2015-2017 data with temperature on a map

```{r}
library(maps)
library(ggplot2)
library(scales)

#load us map data
all_states <- map_data("state")
florida <- all_states[which(all_states$region=="florida"),]

tempNotNA <- which(is.na(bcrs_2015_2017$water_surface_tempature) == FALSE &
                   is.na(bcrs_2015_2017$air_tempature) == FALSE)
bcrs_2015_2017_temp <- bcrs_2015_2017[tempNotNA, ]

ggplot(bcrs_2015_2017_temp, aes(longitude, latitude, color = bcrs_2015_2017_temp$water_surface_tempature)) +
  geom_polygon( data=all_states, aes(x=long, y=lat, group = group),colour="white", fill="grey10" ) +
  #geom_point(alpha = 0.5) + scale_colour_gradientn(colours=rainbow(4)) +
  geom_point(alpha = 0.5, na.rm = TRUE) + #scale_colour_gradient(low = "blue", high = "red") +
  scale_colour_gradient2(low = "blue", mid = "yellow", high = "red", midpoint = 85, space = "Lab",
  na.value = "grey50", guide = "colourbar") +
  coord_cartesian(xlim=c(-90,-80),ylim=c(25,32)) +
  ggtitle(paste("Figure 1\n", "Water Surface Temperature for Beaches")) +
  theme(plot.title = element_text(hjust = 0.5)) +
  guides(colour = guide_colourbar(title="Water Surface Temperature"))
```

```{r}
hist(bcrs_2015_2017_temp$water_surface_tempature)
plot(bcrs_2015_2017_temp$water_surface_tempature,bcrs_2015_2017_temp$air_tempature)
```

There is a cluster of water temperatures that look like they were recorded by Celsius instead of Fahrenheit.  We convert them to Fahrenheit.

```{r}
needConversion <- which(bcrs_2015_2017_temp$water_surface_tempature < 37)
bcrs_2015_2017_temp$water_surface_tempature[needConversion] <-
  bcrs_2015_2017_temp$water_surface_tempature[needConversion] * 1.8 + 32

plot(bcrs_2015_2017_temp$water_surface_tempature,bcrs_2015_2017_temp$air_tempature)
```

Now there is one major cluster of air and water temperatures with three outliers (two for water temp and one for air temp).

```{r}
hist(bcrs_2015_2017_temp$created_at, breaks = "days", freq = TRUE)
```


```{r}
beach1 <- bcrs_2015_2017_temp[which(bcrs_2015_2017_temp$beach_id == 1),]
hist(beach1$created_at, breaks = "days", freq = TRUE)
```

```{r}
beach2 <- bcrs_2015_2017_temp[which(bcrs_2015_2017_temp$beach_id == 2),]
hist(beach2$created_at, breaks = "days", freq = TRUE)
```


```{r}
ggplot(bcrs_2015_2017_temp, aes(longitude, latitude, color = bcrs_2015_2017_temp$water_surface_tempature)) +
  geom_polygon( data=all_states, aes(x=long, y=lat, group = group),colour="white", fill="grey10" ) +
  #geom_point(alpha = 0.5) + scale_colour_gradientn(colours=rainbow(4)) +
  geom_point(alpha = 0.5, na.rm = TRUE) + #scale_colour_gradient(low = "blue", high = "red") +
  scale_colour_gradient2(low = "blue", mid = "yellow", high = "red", midpoint = 85, space = "Lab",
  na.value = "grey50", guide = "colourbar") +
  coord_cartesian(xlim=c(-83,-81.25),ylim=c(26,28.5)) +
  ggtitle(paste("Figure 1\n", "Water Surface Temperature for Beaches")) +
  theme(plot.title = element_text(hjust = 0.5)) +
  guides(colour = guide_colourbar(title="Water Surface Temperature"))


```

# Create Bounding Polygon around points around gulf coast.

```{r}
library(sp)

dataPoly = data.frame(matrix(c(-82.9,28.1,
                  -82.8,28.1,
                  -82.15,26.6,
                  -81.78,26.5,
                  -81.8,26.25,
                  -82.2,26.3,
                  -82.75,27.4,
                  -82.9,28.1), 
             ncol = 2, byrow = TRUE))


P1 = Polygon(dataPoly)
Ps1 = SpatialPolygons(list(Polygons(list(P1), ID = "a")), proj4string=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))
plot(Ps1, axes = TRUE)
```

```{r}
dataPoly = data.frame(matrix(c(1,-82.9,28.1,
                  1,-82.8,28.1,
                  1,-82.15,26.6,
                  1,-81.78,26.5,
                  1,-81.8,26.25,
                  1,-82.2,26.3,
                  1,-82.75,27.4,
                  1,-82.9,28.1), 
             ncol = 3, byrow = TRUE))

colnames(dataPoly) <- c("group","long", "lat")

ggplot(bcrs_2015_2017_temp, aes(longitude, latitude, color = bcrs_2015_2017_temp$water_surface_tempature)) +
  geom_polygon( data=all_states, aes(x=long, y=lat, group = group),colour="white", fill="grey10" ) +
  #geom_point(alpha = 0.5) + scale_colour_gradientn(colours=rainbow(4)) +
  geom_point(alpha = 0.5, na.rm = TRUE) + #scale_colour_gradient(low = "blue", high = "red") +
  scale_colour_gradient2(low = "blue", mid = "yellow", high = "red", midpoint = 85, space = "Lab",
  na.value = "grey50", guide = "colourbar") +
  coord_cartesian(xlim=c(-83,-81.25),ylim=c(26,28.5)) +
  ggtitle(paste("Figure 1\n", "Water Surface Temperature for Beaches")) +
  theme(plot.title = element_text(hjust = 0.5)) +
  guides(colour = guide_colourbar(title="Water Surface Temperature")) +
  geom_polygon(data=dataPoly, aes(x=long, y=lat, group = group),colour="black", fill="grey10", alpha=.5 )

```



Decide how to bin time.

```{r}
bcrs_2015_2017_temp$created_at <- as.POSIXlt(bcrs_2015_2017_temp$created_at)
bcrs_2015_2017_temp$time_bin <- case_when(
  bcrs_2015_2017_temp$created_at$hour < 12 ~ paste(bcrs_2015_2017_temp$created_at$year + 1900,
                                                   bcrs_2015_2017_temp$created_at$mon + 1,
                                                   bcrs_2015_2017_temp$created_at$mday, 
                                                   "AM", sep = "-"),
  TRUE ~ paste(bcrs_2015_2017_temp$created_at$year + 1900,
               bcrs_2015_2017_temp$created_at$mon + 1,
               bcrs_2015_2017_temp$created_at$mday, 
               "PM", sep = "-")
  )

bin1 <- bcrs_2015_2017_temp[which(bcrs_2015_2017_temp$time_bin == "2015-11-9-PM"), ]

ggplot(bin1, aes(longitude, latitude)) + 
  geom_raster(aes(fill = water_surface_tempature),
              interpolate=FALSE) +
  coord_cartesian(xlim=c(-83,-81.25),ylim=c(26,28.5))
```



For a few specific days, look at distribution of unique beaches binned by hour


# Bind the data from the two Excel sheets together.  This will be useful for flags.

```{r}

```




Create animation of point data.

```{r}
library(gganimate)

bcrs_2015_2017_temp_2016 <- bcrs_2015_2017_temp[which(bcrs_2015_2017_temp$created_at$year+1900 == 2016),]

p <- ggplot(bcrs_2015_2017_temp_2015_12, aes(longitude, latitude, color = water_surface_tempature, frame = time_bin)) +
  geom_point(size = 7)

gganimate(p, "C:/Users/lance/Desktop/GIS/Mote_Marine_Lab/output.gif")



# try the example that was in GitHub repository
library(gapminder)
library(ggplot2)
library(animation)

theme_set(theme_bw())

p <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, color = continent, frame = year)) +
  geom_point() +
  scale_x_log10()

magickPath <- shortPathName("C:\\Program Files\\ImageMagick-7.0.6-Q16\\magick.exe")
ani.options(convert=magickPath)

gganimate(p)
```


Strange error:

magick.exe: unable to load module 'C:\Program Files\ImageMagick-7.0.6-Q16\modules\coders\IM_MOD_RL_PNG_.dll': The specified module could not be found.
 @ error/module.c/OpenModule/1266.
magick.exe: no decode delegate for this image format `PNG' @ error/constitute.c/ReadImage/509.




Experiment with interpolation

```{r}
library(sp)
dsp <- SpatialPoints(bin1[,30:29], proj4string=CRS("+proj=longlat +datum=NAD83"))
dsp <- SpatialPointsDataFrame(dsp, d)


```

Create Animations Part 2

```{r}
library(animation)
library(colorRamps)
#changing scale
draw.curve<-function(bin){
  a<-ggplot(bcrs_2015_2017_temp_2016[bcrs_2015_2017_temp_2016$time_bin==bin,],
            aes(longitude,latitude, color = water_surface_tempature)) +
     geom_polygon( data=florida, aes(x=long, y=lat, group = group),colour="white", fill="grey10" ) +
     geom_point(size=7) +
     scale_colour_gradientn(colours = colorRamps::matlab.like(100)) +
     ggtitle(paste("Water Surface Temperature on", bin)) +
     theme(plot.title = element_text(hjust = 0.5))
  
  print(a)
}

#fixed scale
min_water <- min(bcrs_2015_2017_temp_2016$water_surface_tempature)
max_water <- max(bcrs_2015_2017_temp_2016$water_surface_tempature)

draw.curve<-function(bin){
  a<-ggplot(bcrs_2015_2017_temp_2016[bcrs_2015_2017_temp_2016$time_bin==bin,],
            aes(longitude,latitude, color = water_surface_tempature)) +
     geom_polygon( data=florida, aes(x=long, y=lat, group = group),colour="white", fill="grey10" ) +
     geom_point(size=7) +
     scale_colour_gradientn(colours = colorRamps::matlab.like(100), limits=c(min_water, max_water)) +
     ggtitle(paste("Water Surface Temperature on", bin)) +
     theme(plot.title = element_text(hjust = 0.5))
  
  print(a)
}

trace.animate <- function() {
  lapply(unique(bcrs_2015_2017_temp_2016$time_bin), function(i) {
    draw.curve(i)
  })
}

# http://www.wikihow.com/Install-FFmpeg-on-Windows was useful to install ffmpeg
saveVideo(trace.animate(), movie.name = "2016.mp4",
          img.name = "2016_pics",
          ffmpeg = ani.options("C:/ffmpeg/bin/ffmpeg.exe"))
```

```{r}
# flags
draw.curve<-function(bin){
  a<-ggplot(bcrs_2015_2017_temp_2016[bcrs_2015_2017_temp_2016$time_bin==bin,],
            aes(longitude,latitude, color = flag)) +
     geom_polygon( data=florida, aes(x=long, y=lat, group = group),colour="white", fill="grey10" ) +
     geom_point(size=7) +
     scale_colour_manual(values = c("Green" = "green", "Yellow" = "yellow", "Red" = "red", 
                                    "Double Red" = "darkred", "Purple" = "purple")) +
     ggtitle(paste("Flags on", bin)) +
     theme(plot.title = element_text(hjust = 0.5))
  
  print(a)
}

trace.animate <- function() {
  lapply(unique(bcrs_2015_2017_temp_2016$time_bin), function(i) {
    draw.curve(i)
  })
}

saveVideo(trace.animate(), movie.name = "2016.mp4",
          img.name = "2016_pics",
          ffmpeg = ani.options("C:/ffmpeg/bin/ffmpeg.exe"))

```



Add new functionality to the ggplot


```{r}
library(grid)

# add vx and vy columns that will be useful for drawing arrows based on speed and direction
bcrs_2015_2017_temp$vx <- case_when(
  bcrs_2015_2017_temp$wind_direction == "east" ~ bcrs_2015_2017_temp$wind_speed,
  bcrs_2015_2017_temp$wind_direction == "west" ~ -bcrs_2015_2017_temp$wind_speed,
  bcrs_2015_2017_temp$wind_direction == "north" ~ 0,
  bcrs_2015_2017_temp$wind_direction == "south" ~ 0,
  bcrs_2015_2017_temp$wind_direction == "northeast" ~ (bcrs_2015_2017_temp$wind_speed ^ 2 / 2) ^ 0.5,
  bcrs_2015_2017_temp$wind_direction == "northwest" ~ -(bcrs_2015_2017_temp$wind_speed ^ 2 / 2) ^ 0.5,
  bcrs_2015_2017_temp$wind_direction == "southeast" ~ (bcrs_2015_2017_temp$wind_speed ^ 2 / 2) ^ 0.5,
  bcrs_2015_2017_temp$wind_direction == "southwest" ~ -(bcrs_2015_2017_temp$wind_speed ^ 2 / 2) ^ 0.5
)

bcrs_2015_2017_temp$vy <- case_when(
  bcrs_2015_2017_temp$wind_direction == "east" ~ 0,
  bcrs_2015_2017_temp$wind_direction == "west" ~ 0,
  bcrs_2015_2017_temp$wind_direction == "north" ~ bcrs_2015_2017_temp$wind_speed,
  bcrs_2015_2017_temp$wind_direction == "south" ~ -bcrs_2015_2017_temp$wind_speed,
  bcrs_2015_2017_temp$wind_direction == "northeast" ~ (bcrs_2015_2017_temp$wind_speed ^ 2 / 2) ^ 0.5,
  bcrs_2015_2017_temp$wind_direction == "northwest" ~ (bcrs_2015_2017_temp$wind_speed ^ 2 / 2) ^ 0.5,
  bcrs_2015_2017_temp$wind_direction == "southeast" ~ -(bcrs_2015_2017_temp$wind_speed ^ 2 / 2) ^ 0.5,
  bcrs_2015_2017_temp$wind_direction == "southwest" ~ -(bcrs_2015_2017_temp$wind_speed ^ 2 / 2) ^ 0.5
)

bcrs_2015_2017_temp$vx <- bcrs_2015_2017_temp$vx * .2 
bcrs_2015_2017_temp$vy <- bcrs_2015_2017_temp$vy * .2 

bcrs_2015_2017_temp$resp_color <- case_when(
  bcrs_2015_2017_temp$respiratory_irritation == "none" ~ "green",
  bcrs_2015_2017_temp$respiratory_irritation == "slight" ~ "yellow",
  bcrs_2015_2017_temp$respiratory_irritation == "moderate" ~ "orange",
  bcrs_2015_2017_temp$respiratory_irritation == "intense" ~ "red"
)

bcrs_2015_2017_temp_2016 <- bcrs_2015_2017_temp[which(bcrs_2015_2017_temp$created_at$year+1900 == 2016),]

# practice on one plot for first time bin
bin = unique(bcrs_2015_2017_temp_2016$time_bin)[2]
df <- bcrs_2015_2017_temp_2016[bcrs_2015_2017_temp_2016$time_bin==bin,]
ggplot(df) +
  geom_polygon( data=florida, aes(x=long, y=lat, group = group),colour="white", fill="grey10" ) +
  geom_point(size = 7, shape=21, aes(longitude,latitude, fill = water_surface_tempature)) +
  scale_fill_gradientn(colours = colorRamps::matlab.like(100), limits=c(min_water, max_water)) +
  geom_segment(mapping = aes(x=longitude, y=latitude, xend=longitude+vx, yend=latitude+vy, color=respiratory_irritation),
               arrow = arrow(angle=15), size = 1) +
  scale_colour_manual(values = c("none" = "green", "slight" = "yellow", "moderate" = "orange", "intense" = "red")) + 
  ggtitle(paste("Respiratory Irritation", bin)) +
  theme(plot.title = element_text(hjust = 0.5))


```

```{r}
draw.curve<-function(bin){
  df <- bcrs_2015_2017_temp_2016[bcrs_2015_2017_temp_2016$time_bin==bin,]
  a <- ggplot(df) +
    geom_polygon( data=florida, aes(x=long, y=lat, group = group),colour="white", fill="grey10" ) +
    geom_point(size = 7, shape=21, aes(longitude,latitude, fill = water_surface_tempature)) +
    scale_fill_gradientn(colours = colorRamps::matlab.like(100), limits=c(min_water, max_water)) +
    geom_segment(mapping = aes(x=longitude, y=latitude, xend=longitude+vx, yend=latitude+vy,
                               color=respiratory_irritation),
                 arrow = arrow(angle=15), size = 1) +
    scale_colour_manual(values = c("none" = "green", "slight" = "yellow", "moderate" = "orange", "intense" = "red")) + 
    ggtitle(paste("Respiratory Irritation", bin)) +
    theme(plot.title = element_text(hjust = 0.5))
  
  print(a)
}

trace.animate <- function() {
  lapply(unique(bcrs_2015_2017_temp_2016$time_bin), function(i) {
    draw.curve(i)
  })
}

saveVideo(trace.animate(), movie.name = "2016.mp4",
          img.name = "2016_pics",
          ffmpeg = ani.options("C:/ffmpeg/bin/ffmpeg.exe"))
```

